\sysfull

### 2018-04-16 paper/paging.tex
paging

\sys exposes non-volatile memory for \emph{protected} variables to
the application as a virtual address space partitioned into pages.
Paged virtual memory is a key enabler of task coalescing, which
we described in Section~\ref{sec:task_coalescing}. Virtualization simplifies
coalescing, because the system gets the freedom to move application's data
between volatile and non-volatile memory transparently.
%
\sys manages memory such that application tasks operate primarily from
efficient volatile SRAM rather than non-volatile FRAM, which is more costly to access than SRAM,
as we showed in Figure~\ref{fig:framEnergy}. \sys runtime implements memory
virtualization by moving pages of the address space from the large,
non-volatile memory to a small, volatile memory on demand and writing all dirty
pages back to non-volatile memory atomically. Since memory moves at page granularity, \sys
exploits hardware\hyp{}accelerated block copies using DMA.

@@

Paging moves data between non-volatile and volatile memory. A page has a home location in non-volatile memory which we refer to as the page's \texttt{\underline{store}} (underlined locations are non-volatile).
As a task accesses pages, they are swapped from their home location
into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The \texttt{working} buffer is allocated to take up the volatile memory that remains after space for stack is allocated.
%
\sys uses a two-phase commit to write dirty pages from the \texttt{working} buffer back to their home location in \texttt{\underline{store}} atomically, ensuring data consistency even with power failures.
%
When any dirty page is swapped out from the \texttt{working} buffer due to eviction or task transition from last coalesced task, the first phase writes the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer, and the second phase writes it from \texttt{\underline{shadow}} to \texttt{\underline{store}}.
%
Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a time between volatile and non-volatile memory, \sys's page-based design moves contiguous blocks of data.
%
\sys takes full advantage of hardware\hyp{}accelerated block copies using Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x fewer cycles on MSP430FRxxxx.

>>>>

\sys exposes non-volatile memory for \emph{protected} variables to
the application as a virtual address space partitioned into pages.
\emph{Paged virtual memory} is a key enabler of task coalescing since it keeps the non-volatile memory consistent despite dynamic WAR dependencies incurred by coalesced code regions. Thanks to memory virtualization, \sys have the freedom to move application's data between volatile and non-volatile memory transparently.
%
%,. Virtualization simplifies
%coalescing, because the system gets the freedom to move application's data
%between volatile and non-volatile memory transparently.
%
To this end, \sys isolates FRAM from the tasks of the application so that they interface only SRAM---which is cheaper to read/write as we indicated in Section~\ref{sec:background}.
%
%manages memory such that application tasks operate primarily from
%efficient volatile SRAM rather than non-volatile FRAM, which is more costly to access than SRAM,
%as we showed in Figure~\ref{fig:framEnergy}. 
\sys runtime implements memory
virtualization by moving pages of the address space from the large,
non-volatile memory to a small, volatile memory on demand and writing all dirty
pages back to non-volatile memory atomically. Since memory moves at page granularity, \sys
exploits hardware\hyp{}accelerated block copies using DMA.

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
	\caption{Illustration of task-virtual memory interaction in \sys. Tasks only interact with the working buffer in SRAM (indicated as page buffer). The shadow buffer is used to hold the dirty bages. The indirection table is used to speed-up page two-phase commit operation: it eliminates the copy operation in the second phase.}
	\label{figure:coala_page_tags}
\end{figure}

\textbf{System Overview.} The overall picture of the paging operations is illustrated in Figure~\ref{figure:coala_page_tags}. Paging moves data between non-volatile and volatile memory. A page has a home location in non-volatile memory which we refer to as the page's \texttt{\underline{store}} (underlined locations are non-volatile).
As a task accesses pages, they are swapped from their home location
into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The \texttt{working} buffer is allocated to take up the volatile memory that remains after space for stack is allocated.
%
\sys uses a two-phase commit to write dirty pages from the \texttt{working} buffer back to their home location in \texttt{\underline{store}} atomically, ensuring data consistency even with power failures.
%
When any dirty page is swapped out from the \texttt{working} buffer due to eviction or task transition from last coalesced task, the first phase writes the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer, and the second phase writes it from \texttt{\underline{shadow}} to \texttt{\underline{store}}.
%
Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a time between volatile and non-volatile memory, \sys's page-based design moves contiguous blocks of data.
%
\sys takes full advantage of hardware\hyp{}accelerated block copies using Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x fewer cycles on MSP430FRxxxx.
